---
title: "gap_final"
author: "Hadil"
date: "2023-02-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
install.packages("optimx")
install.packages("lme4")
install.packages("Matrix")
install.packages("ggplot2")
install.packages("usethis")
install.packages("tidyverse")
install.packages("nlme")
install.packages("janitor")
install.packages("emmeans")
install.packages("sjstats")
install.packages("lmerTest")
install.packages("MuMIn")
devtools::install_github('cttobin/ggthemr')
devtools::install_github("hadley/devtools")


```


```{r}
library(Matrix)
library(readxl)
library(usethis)
library(ggplot2)
library(ggthemr)
library(lme4)
library(devtools)
library(Matrix)
library(DT)
library(optimx)
library(ggdist)
library(brms)
library(ggeffects)
library(dplyr)
library(tidyverse)
library(nlme)
library(emmeans)
library(lmerTest)
```


```{r}
ggthemr("fresh")
library(readxl)
Gap_final <- read_excel("Gap_final.xlsx")
View(Gap_final)
  janitor::clean_names() |>
  mutate(participant = as.factor(participant),
         group = as.factor(group),
  )

knitr::opts_chunk$set(echo = TRUE)
```


```{r}
hist(Gap_final$RT)
log_rt <- log(Gap_final$RT)
hist(log_rt)
```

## coding level factors
```{r}
Gap_final$Condition = ifelse(Gap_final$condition == "CONG", -0.5, 0.5)
```

```{r}
Gap_final <- Gap_final %>%
  mutate(hs1 = case_when(group == "AHS" ~ -1L, # baseline
                          group == "HS1" ~ 1L,  # target
                          TRUE      ~ 0L), # anything else
         hs2 = case_when(group == "AHS" ~ -1L, # baseline
                          group == "HS2" ~  1L, # target
                          TRUE      ~ 0L), # anything else
          rt1 = case_when(group == "AHS" ~ -1L, # baseline
                          group == "RT1" ~  1L, # target
                          TRUE      ~ 0L), # anything else
          rt2 = case_when(group == "AHS" ~ -1L, # baseline
                          group == "RT2" ~  1L, # target
                          TRUE      ~ 0L))

```


```{r}
Gap_final$group <- as.factor(Gap_final$group)
Gap_final$item <- as.factor(Gap_final$item)
Gap_final$key_resp_5_corr <- as.factor(Gap_final$key_resp_5_corr)

str(Gap_final)
```



# RT per condition
```{r}
Gap_final |>
  ggplot(aes(x = condition, y = log_rt, fill = group)) +
  geom_boxplot()
```
```{r}
Gap_final |>
  mutate(key_resp_5_corr = as.numeric(as.character(key_resp_5_corr))) %>%
  ggplot(aes(x = condition, y = key_resp_5_corr, fill = group)) +
  geom_boxplot()
```


```{r}
library(dplyr)

summary_table <- Gap_final %>%
  group_by(group, Condition) %>%
  summarise(across(where(is.numeric),
    Mean = mean(log_rt),
    SD = sd(log_rt)
  ) %>%
  ungroup()

print(summary_table)
```


## Data summary

```{r}
Gap_final %>%
group_by(group) %>%
  group_by(condition, group) %>%
  summarise(across(where(is.numeric),
    list(mean = ~ mean(.x, na.rm = TRUE), count = ~ sum(!is.na(.x)))))
```
```{r}
Gap_final %>%
  group_by(group) %>%
  group_by(condition, group) %>%
  summarise(across(
    where(is.numeric),
    list(mean = ~mean(.x, na.rm = TRUE),
         count = ~sum(!is.na(.x)),
         sd = ~sd(.x, na.rm = TRUE),     # Add standard deviation calculation
         range = ~diff(range(.x, na.rm = TRUE)))   # Add range calculation
    ))
```
```{r}
Gap_final %>%
  group_by(group) %>%
  group_by(condition, group) %>%
  summarise(across(
    where(is.numeric),
    list(mean = ~mean(.x, na.rm = TRUE),
         count = ~sum(!is.na(.x)),
         sd = ~sd(.x, na.rm = TRUE),
         range = ~diff(range(.x, na.rm = TRUE)),
         min = ~min(.x, na.rm = TRUE),     # Add minimum calculation
         max = ~max(.x, na.rm = TRUE)      # Add maximum calculation
    ))) 
```


```{r}
Gap_final %>%
group_by(group) %>%
  summarise(across(where(is.numeric),
    list(mean = ~ mean(.x, na.rm = TRUE), count = ~ sum(!is.na(.x)))))

```
```{r}
Gap_final %>%
  group_by(group) %>%
  summarise(across(
    where(is.numeric),
    list(mean = ~mean(.x, na.rm = TRUE),
         count = ~sum(!is.na(.x)),
         sd = ~sd(.x, na.rm = TRUE),     # Add standard deviation calculation
         range = ~diff(range(.x, na.rm = TRUE)))   # Add range calculation
  ))
```

```{r}
Gap_final %>%
  group_by(group) %>%
  summarise(across(
    where(is.numeric),
    list(mean = ~mean(.x, na.rm = TRUE),
         count = ~sum(!is.na(.x)),
         sd = ~sd(.x, na.rm = TRUE),     # Add standard deviation calculation
         range = ~diff(range(.x, na.rm = TRUE)),   # Add range calculation
         min = ~min(.x, na.rm = TRUE),     # Add minimum calculation
         max = ~max(.x, na.rm = TRUE))     # Add maximum calculation
  ))
```



```{r}
library(dplyr)

sum_table <- Gap_final %>%
  mutate(key_resp_5_corr = as.numeric(as.character(key_resp_5_corr))) %>%
  group_by(condition, group) %>%
  summarise(
    Total = mean(key_resp_5_corr)
  ) %>%
  ungroup()

print(sum_table)
```
```{r}
sum_table <- Gap_final %>%
  mutate(key_resp_5_corr = as.numeric(as.character(key_resp_5_corr))) %>%
  group_by(condition, group) %>%
  summarise(
    Total = mean(key_resp_5_corr),
    SD = sd(key_resp_5_corr),
    Min = min(key_resp_5_corr),
    Max = max(key_resp_5_corr)
  ) %>%
  ungroup()

print(sum_table)
```



```{r}


#first load in the packages you want
require(lme4)
require(MuMIn)
install.packages("rsq")
library(rsq)
```

```{r}
# Fit a model including fixed and all random effects
mF <- lmer(log_rt ~ group + length + condition + group * condition +group * Dom_scale + group*English_scale + group*Arabic_scale + (1|participant) + (1|item), data = Gap_final)

mF2 <- glmer(key_resp_5_corr ~ group + condition+ group * condition + (1|participant) + (1|item),
               family = binomial(link = "logit"), 
               data = Gap_final,
               nAGQ=settingAGQ,control = lbfgs_ctrl)

anova(mF)
eta_sq(mF, partial = TRUE)
r.squaredGLMM(mF)
r.squaredGLMM(mF2)
emmeans (mF, list( pairwise ~ group), adjust = "tukey")
qqnorm(resid(mF))
effectsize::eta_squared(mF)
effectsize::eta_squared(mF2)
rsq.lmm(mF,adj=FALSE)

```

```{r}
Gap_final |>
  ggplot(aes(x = group, y = key_resp_5_corr, fill = Condition)) +
  geom_boxplot()
```

```{r}
library(data.table)
freqs <- Gap_final |>
  group_by(group, condition) |>
  summarise(total_correct = sum(key_resp_5_corr),
            fract_correct = total_correct / n()) |>
  arrange(condition, group)

freqs |> data.table()
```
# Correct answers per group & condition, visualized as bar plot
## i couldn't run this, showed me an error.

```{r, fig.width=20, fig.height=6}
freqs |>
  ggplot(aes(x = group, y =key_resp_5_corr, fill = group)) +
  geom_col() +
  facet_wrap(~ condition, ncol = 4) +
  theme(legend.position = "bottom")
```

# RTs per group & condition  ( also error )

```{r}
rts <- Gap_final|>
  filter(!is.na(log_rt)) |>
  group_by(group, condition) |> 
  summarise(median_rt = median(log_rt),
            iqr_rt = IQR(log_rt)) |> 
  arrange(condition, group)

rts |> data.table()
```

# RTs visualized using boxplots

```{r, fig.width=20, fig.height=6}

Gap_final |>
  ggplot(aes(x = group, y = log_rt, fill = condition)) +
  geom_boxplot() +
  facet_wrap(~ condition, ncol = 4) +
  theme(legend.position = "bottom")

```


# Anomalous participant: unusually large number of observations

```{r}
Gap_final|> 
  group_by(participant, condition) |> 
  count() |> 
  pull(n) |> 
  map_lgl(\(x) x == 23) |> 
  all()
```

# Models (Accuracy) (Binomial Logistic Regression)
#I first ran an intercept only model, with correct answers as the dependent variable, with  a random intercept for participants. Mod_0 does not include a fixed effect for group.
#In Mod_2 I added group 
#in mod 3 i added participant and item as random intercepts
#AIC and BIC lower for mod_3 and significant improvement over model-0
#Therefore, mod_3 is a better fit  for the data.
## AHS > RT2 > HS1 > RT1
## AHS = HS2

## but with post hoc it showed no significant difference between AHS and RT2. only with mixed effect it did show.
# HS2 = RT2 = AHS

```{r}
# mod0 <- glmer(as.factor(key_resp_5.CORR) ~ group + (1|group/participant) + (1|condition), family = binomial, data = GapTaskLong, control = glmerControl(check.conv.grad = .makeCC("warning", tol = 5e-2)))

lbfgs_ctrl <- glmerControl()

xls_nb <- Gap_final |>
  mutate(condition = as.factor(condition) |> relevel(ref = "CONG"),
         group = as.factor(group) |> relevel(ref = "AHS"),
         correct = key_resp_5_corr == 1)

mod_0 <- glmer(key_resp_5_corr ~ 1 + (1|participant),
               family = binomial(link = "logit"), 
               data = Gap_final,
               control = lbfgs_ctrl)

mod_1 <- glmer(key_resp_5_corr ~ 1 + (1|participant) + (1|item),
               family = binomial(link = "logit"), 
               data = Gap_final,
               control = lbfgs_ctrl)

mod_2 <- glmer(key_resp_5_corr ~ group + (1|participant),
               family = binomial(link = "logit"),
               data = Gap_final,
               control = lbfgs_ctrl)

mod_3 <- glmer(key_resp_5_corr ~ group + (1|participant) + (1|item),
               family = binomial(link = "logit"),
               data = Gap_final,
               control = glmerControl(optimizer = "bobyqa"))

mod_4 <- glmer(key_resp_5_corr ~ group + (hs1 + hs2 + rt1 + rt2) * condition + (1|participant) + (1|item), family = binomial(link = "logit"), data = Gap_final, nAGQ=settingAGQ,control = lbfgs_ctrl)

mod_5 <- glmer(key_resp_5_corr ~ group + (hs1 + hs2 + rt1 + rt2) * condition + (1|participant) + (1|item) +(1+condition|participant), family = binomial(link = "logit"), data = Gap_final, nAGQ=settingAGQ,control = lbfgs_ctrl)



mod_6 <- glmer(key_resp_5_corr ~ group + (hs1 + hs2 + rt1 + rt2) * condition + (hs1 + hs2 + rt1 + rt2) * Arabic_scale + (hs1 + hs2 + rt1 + rt2) * English_scale + (hs1 + hs2 + rt1 + rt2) * Dom_scale + (1|participant) + (1|item), family = binomial(link = "logit"), data = Gap_final, nAGQ=settingAGQ,control = lbfgs_ctrl)

mod_7 <- glmer(key_resp_5_corr ~ group + (hs1 + hs2 + rt1 + rt2) * condition + (hs1 + hs2 + rt1 + rt2) * Arabic_scale + (hs1 + hs2 + rt1 + rt2) * English_scale + (hs1 + hs2 + rt1 + rt2) * Dom_scale + (hs1 + hs2 + rt1 + rt2) * English_att_scale + (hs1 + hs2 + rt1 + rt2) * Arabic_att_scale + (1|participant) + (1|item), family = binomial(link = "logit"), data = Gap_final, nAGQ=settingAGQ,control = lbfgs_ctrl)




print(anova(mod_0, mod_1, mod_2, mod_3, mod_4, mod_5, mod_6, mod_7))
```

```{r}
library(lmerTest)
print(summary(mod_4))
anova(mod_3)
```
##Model estimates (Accuracy with group)
This clearly indicates a significant main effect is shown between the returnees and the AHS.

```{r}
model_params <- broom.mixed::tidy(mod_4)

model_params|>
  filter(!str_detect(term, "sd")) |>
  ggplot(aes(x = term, y = estimate)) +
  # geom_hline(yintercept = 1, linetype = "dotted", color = "black") +
  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), width = 0.1, color = "black")
  geom_point(aes(color = term), size = 3) +
  theme(legend.position = "bottom") +
  labs(y = "Accuracy", x = "Group")
```
```{r}
model_params <- broom.mixed::tidy(mod_4)
model_params |>
  filter(!str_detect(term, "sd")) |>
  ggplot(aes(x = term, y = estimate, color = term)) +
  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), width = 0.3, color = "black") +
  geom_point(size = 4) +
  scale_color_manual(values = c("red", "blue", "green", "orange", "purple", "pink", "yellow", "maroon", "brown", "black")) +  # Adjust the colors as needed
  theme(legend.position = "bottom")+
  labs(y = "Accuracy", x = "Group")
```
```{r}
library(lmerTest)
print(summary(mod_7))
anova(mod_7)
```

# Here I attempt to perform post-hoc by using emmeans for mod_3 (accuracy)

```{r}
library(emmeans)
emmeans(mod_3, list(pairwise ~ group), adjust = "tukey")
```
```{r}
library(ggplot2)
ggplot(Gap_final, aes(x = group, y = log_rt, color = group)) +
  geom_point() +
  geom_smooth(method='lm',formula=y~x)
```

```{r}
sjPlot::plot_model(mod_3)
```


# Models with RT ( when using log rt only without log residuals rts showed an effect)

# RTS were log-transformed here only

## AHS = HS2 = RT2
## with length only a significant shown among HS2 against AHS
# post hoc shows RT1 = HS1 and RT > RT1 and  AHS = HS2 = RT2

```{r}
mod0_rt <- lmer(log_rt ~ 1 + (1|participant), data = Gap_final)

mod1_rt <- lmer(log_rt ~ 1 + (1|participant) + (1|item), data = Gap_final)

mod2_rt <- lmer(log_rt ~ group + (1|participant), data = Gap_final)

mod3_rt <- lmer(log_rt ~ group + length + (1|participant) + (1|item), data = Gap_final)

mod4_rt <- lmer(log_rt ~ group + length + group*length + (1|participant) + (1|item) , data = Gap_final)



print(anova(mod0_rt, mod1_rt, mod2_rt, mod3_rt, mod4_rt))
```

```{r}
summary(mod3_rt)
```


```{r}
ranef(mod4_rt)
hist(ranef(mod4_rt)$participant[,1])
hist(ranef(mod4_rt)$item[,1])
hist(resid(mod4_rt))
plot(mod5_rt)
```




# In the graph below it clearly shows a significant main effect between the AHS and the other groups.

```{r}
model_params_rt <- broom.mixed::tidy(mod3_rt)

model_params_rt |> 
  filter(!str_detect(term, "sd")) |>
  ggplot(aes(x = term, y = estimate)) +
  # geom_hline(yintercept = 1, linetype = "dotted", color = "black") +
  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), width = 0.1, color = "black") +
  geom_point(aes(color = term), size = 3) +
  theme(legend.position = "bottom") +
  labs(y = "RT", x = "Group")
```

# Here I attempt to perform post-hoc by using emmeans for  (reaction time) only

```{r}
library(emmeans)
emmeans(mod4_rt, list(pairwise ~ group), adjust = "bonferroni")
```
#Models with interaction with condition for ACCURACY

```{r}
settingAGQ=0
```



```{r}
str(Gap_final)
summary (Gap_final$key_resp_5_corr)
```


## interaction between group and condition for ACCURACY
##condition and group as a fixed effects and a random intercept for participants and for items. “congruent” condition was the reference level for condition.

#model 5 better fit
adding length shows response is constant in the ldt task it worked

## A significant effect among HS1 of Incongruent trials only against the AHS

```{r}
Gap_final$Dominance_score <- scale(Gap_final$Dominance_score, center = TRUE, scale = TRUE)
Gap_final$English_use <- scale(Gap_final$English_use, center = TRUE, scale = TRUE)
Gap_final$Arabic_use <- scale(Gap_final$Arabic_use, center = TRUE, scale = TRUE)
colnames(Gap_final)[35] <- "Dom_scale"
colnames(Gap_final)[27] <- "English_scale"
colnames(Gap_final)[28] <- "Arabic_scale"
Gap_final$English_attidutes <- scale(Gap_final$English_attidutes, center = TRUE, scale = TRUE)
Gap_final$Arabic_attidutes <- scale(Gap_final$Arabic_attidutes, center = TRUE, scale = TRUE)
Gap_final$Arabic_history <- scale(Gap_final$Arabic_history, center = TRUE, scale = TRUE)
Gap_final$English_history <- scale(Gap_final$English_history, center = TRUE, scale = TRUE)
colnames(Gap_final)[31] <- "English_att_scale"
colnames(Gap_final)[32] <- "Arabic_att_scale"
colnames(Gap_final)[25] <- "Eng_hist_scale"
colnames(Gap_final)[26] <- "Arab_hist_scale"



```



```{r}



mod_0cd <- glmer(key_resp_5_corr ~ 1 + (1|participant), 
               family = binomial(link = "logit"), 
               data = Gap_final,
             nAGQ=settingAGQ,control = lbfgs_ctrl)

mod_1cd <- glmer(key_resp_5_corr ~ 1 + (1|participant) + (1|item), 
               family = binomial(link = "logit"), 
               data = Gap_final,
               nAGQ=settingAGQ,control= lbfgs_ctrl)

mod_2cd <- glmer(key_resp_5_corr ~ group + (1|participant), 
               family = binomial(link = "logit"), 
               data = Gap_final,
              nAGQ=settingAGQ,control = lbfgs_ctrl)

mod_3cd <- glmer(key_resp_5_corr ~ group + (1|participant) + (1|item), 
               family = binomial(link = "logit"), 
               data = Gap_final,
             nAGQ=settingAGQ,control  = lbfgs_ctrl)

mod_4cd <- glmer(key_resp_5_corr ~ group * condition + (1|participant),
               family = binomial(link = "logit"), 
               data = Gap_final,
               nAGQ=settingAGQ,control = lbfgs_ctrl)

mod_5cd <- glmer(key_resp_5_corr ~ hs1+hs2+rt1+rt2 + Condition+ hs1+hs2+rt1+rt2 * Condition + (1|participant) + (1|item),
               family = binomial(link = "logit"), 
               data = Gap_final,
               nAGQ=settingAGQ,control = lbfgs_ctrl)

mod_6cd <- glmer(key_resp_5_corr ~ hs1+hs2+rt1+rt2*Dom_scale + (1|participant) + (1|item),
               family = binomial(link = "logit"), 
               data = Gap_final,
               nAGQ=settingAGQ,control = lbfgs_ctrl)

mod_7cd <- glmer(key_resp_5_corr ~ hs1+hs2+rt1+rt2 + Condition + hs1+hs2+rt1+rt2*Condition + hs1+hs2+rt1+rt2*Dom_scale + hs1+hs2+rt1+rt2*English_scale + hs1+hs2+rt1+rt2*Arabic_scale + (1|participant) + (1|item) + (1+Condition|participant),
               family = binomial(link = "logit"), 
               data = Gap_final,
               nAGQ=settingAGQ,control = lbfgs_ctrl)

mod_8cd <- glmer(key_resp_5_corr ~ hs1+hs2+rt1+rt2 + Condition + hs1+hs2+rt1+rt2*Condition + Dom_scale + English_scale + Arabic_scale + (1|participant) + (1|item),
              family = binomial(link = "logit"), 
               data = Gap_final,
               nAGQ=settingAGQ,control = lbfgs_ctrl)

mod_9cd <- glmer(key_resp_5_corr ~ hs1+hs2+rt1+rt2 + Condition + hs1+hs2+rt1+rt2*Condition + Dom_scale + English_scale + Arabic_scale + English_att_scale + Arabic_att_scale + Arab_hist_scale + Eng_hist_scale + (1|participant) + (1|item),
              family = binomial(link = "logit"), 
               data = Gap_final,
               nAGQ=settingAGQ,control = lbfgs_ctrl)

mod_10cd <- glmer(key_resp_5_corr ~ hs1+hs2+rt1+rt2 + Condition + Dom_scale + English_scale + Arabic_scale + hs1+hs2+rt1+rt2*Condition + hs1+hs2+rt1+rt2*Dom_scale + hs1+hs2+rt1+rt2*English_scale + hs1+hs2+rt1+rt2*Arabic_scale + (1|participant) + (1|item) + (1+Condition|participant),
               family = binomial(link = "logit"), 
               data = Gap_final,
               nAGQ=settingAGQ,control = lbfgs_ctrl)

mod_11cd <- glmer(key_resp_5_corr ~ hs1+hs2+rt1+rt2 + Condition + Dom_scale + English_scale + Arabic_scale + hs1*Condition + hs2*Condition+ rt1*Condition + rt2*Condition + hs1+hs2+rt1+rt2*Dom_scale + (1|participant) + (1|item),
               family = binomial(link = "logit"), 
               data = Gap_final,
               nAGQ=settingAGQ,control = lbfgs_ctrl)

mod_12cd <- glmer(key_resp_5_corr ~ group + (hs1 + hs2 + rt1 + rt2) * condition + (hs1 + hs2 + rt1 + rt2) * Arabic_scale + (hs1 + hs2 + rt1 + rt2) * English_scale + (1|participant) + (1|item), family = binomial(link = "logit"), data = Gap_final, nAGQ=settingAGQ,control = lbfgs_ctrl)

mod_13cd <- glmer(key_resp_5_corr ~ (hs1 + hs2 + rt1 + rt2) * condition + (1|participant) + (1|item), family = binomial(link = "logit"), data = Gap_final, nAGQ=settingAGQ,control = lbfgs_ctrl)

mod_14cd <- glmer(key_resp_5_corr ~ group * condition + (1|participant)  + (1|item),
               family = binomial(link = "logit"), 
               data = Gap_final,
               nAGQ=settingAGQ,control = lbfgs_ctrl)

mod_15cd <- glmer(key_resp_5_corr ~ group * Condition + (1|participant)  + (1|item),
               family = binomial(link = "logit"), 
               data = Gap_final,
               nAGQ=settingAGQ,control = lbfgs_ctrl)





print(anova(mod_0cd, mod_1cd, mod_2cd, mod_3cd, mod_4cd, mod_5cd,mod_6cd,mod_7cd,mod_8cd,mod_9cd,mod_10cd,mod_11cd, mod_12cd, mod_13cd, mod_14cd, mod_15cd))
```

```{r}
summary(mod_15cd)
```

```{r}
summary(mod_9cd)
```


```{r}
# Generate sample data
set.seed(123)
key_resp_5_corr <- as.character(runif(100))  # Variable 1 (sample data)
English_scale <- as.character(runif(100))  # Variable 2 (sample data)
group <- rep(c("RT1", "RT2"), each = 50)  # Grouping factor

# Create a data frame
Gap_final <- data.frame(key_resp_5_corr, English_scale, group)

# Specify the group for correlation and plotting
group_for_plot <- "RT1"

# Filter the data for the specified group
filtered_data <- Gap_final[Gap_final$group == group_for_plot, ]

# Convert variables to numeric
filtered_data$key_resp_5_corr <- as.numeric(filtered_data$key_resp_5_corr)
filtered_data$English_scale <- as.numeric(filtered_data$English_scale)

# Calculate the correlation within the specified group
correlation <- cor(filtered_data$key_resp_5_corr, filtered_data$English_scale, use = "complete.obs")

print(correlation)

# Create a scatter plot for the specified group
ggplot(filtered_data, aes(x = key_resp_5_corr, y = English_scale, color = group)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  xlab("ACCURACY") +
  ylab("ENGLISH USE") +
  ggtitle(paste("Scatter Plot:", group_for_plot, "(Correlation:", round(correlation, 2), ")"))
```


```{r}
ggplot(Gap_final, aes(x = key_resp_5_corr, y =English_scale, color= group)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  xlab("accuracy") +
  ylab("english use") +
  ggtitle("Scatter Plot: accuracy vs english use")
```

## attempt to apply a correlation between the model of accuracy of the gap task with the specific variable (english use) which showed a weak positive correlation (0.001) 
```{r}
residuals <- residuals(mod_15cd)
English_scale <- Gap_final$English_scale
correlation_english_use <- cor(residuals, English_scale)
print(correlation_english_use)
```
```{r}
# Extract the residuals from the model
residualsGroupRT1 <- residuals(mod_15cd)

# Specify the specific group and variable for correlation
specific_group <- "RT1"
English_scale <- Gap_final$English_scale

# Create a subset of data for the specific group
subset_data <- Gap_final[Gap_final$group == specific_group, ]

# Match the residuals with the subset_data using a common identifier
matched_residuals <- residualsGroupRT1[match(subset_data$common_identifier, Gap_final$common_identifier)]

# Convert variables to numeric if necessary
matched_residuals <- as.numeric(as.character(matched_residuals))
subset_data$English_scale <- as.numeric(as.character(subset_data$English_scale))

# Calculate the correlation between residuals and the variable of interest within the specific group
correlation <- cor(matched_residuals, subset_data$English_scale)
# Print the correlation value
print(correlation)
```

```{r}

```


```{r}
df <- data.frame(residuals = residuals, English_scale= English_scale)
ggplot(df, aes(x = English_scale, y = residuals)) +
  geom_point() +
  xlab("English_use") +
  ylab("Residuals") +
  ggtitle("Scatter Plot: Residuals vs English use")
```

## attempt to apply a correlation between the model of accuracy of the gap task with the specific variable (arabic use) which shows a negative correlation -0.001
```{r}
residuals <- residuals(mod_15cd)
Arabic_scale <- Gap_final$Arabic_scale 
correlation_arabic_use <- cor(residuals, Arabic_scale)
print(correlation_arabic_use)
```
## attempt to apply a correlation between the model of accuracy of the gap task with the specific variable "english attitudes" which shows a negative correlation -0.001
```{r}
residuals <- residuals(mod_15cd)
English_att_scale <- Gap_final$English_att_scale
correlation_english_att <- cor(residuals, English_att_scale)
print(correlation_english_att)
```
## attempt to apply a correlation between the model of accuracy of the gap task with the specific variable "arabic attitudes" which shows a negative correlation -0.000
```{r}
residuals <- residuals(mod_15cd)
Arabic_att_scale <- Gap_final$Arabic_att_scale
correlation_arabic_att <- cor(residuals, Arabic_att_scale)
print(correlation_arabic_att)
```
## attempt to apply a correlation between the model of accuracy of the gap task with the specific variable "english global score" which shows a negative correlation -0.006
```{r}
residuals <- residuals(mod_15cd)
English_global <- Gap_final$English_global
correlation_english_global <- cor(residuals, English_global)
print(correlation_english_global)
```
## attempt to apply a correlation between the model of accuracy of the gap task with the specific variable "arabic global score" which shows a negative correlation -0.000
```{r}
residuals <- residuals(mod_15cd)
Arabic_global <- Gap_final$Arabic_global
correlation_arabic_global <- cor(residuals, Arabic_global)
print(correlation_arabic_global)
```

```{r}
str(Gap_final)
```


```{r}
Gap_final$group <- as.factor(Gap_final$group)
Gap_final$item <- as.factor(Gap_final$item)
Gap_final$condition <- as.factor(Gap_final$condition)
Gap_final$key_resp_5_corr <- as.factor(Gap_final$key_resp_5_corr)

str(Gap_final)
```


```{r}
print(summary(mod_5cd))
anova(mod_5cd)
```
```{r}
library(ggplot2)

# Creating example data
set.seed(123)
df <- data.frame(
  group = rep(c("AHS", "HS2"), each = 50),
  condition = rep(c("congruent", "incongruent"), times = 50),
  outcome = rnorm(100, mean = 10, sd = 2)
)

# Creating the interaction plot
ggplot(df, aes(x = condition, y = outcome, color = group, group = group)) +
  geom_point() +
  geom_line() +
  labs(x = "Condition", y = "Outcome") +
  scale_color_manual(values = c("blue", "red"),
                     labels = c("AHS", "HS2")) +
  theme_minimal()
```
```{r}
# Creating example data
set.seed(123)
df <- data.frame(
  group = rep(c("AHS", "RT2"), each = 50),
  condition = rep(c("CONG"), times = 50),
  outcome = rnorm(100, mean = 10, sd = 2)
)

# Creating the interaction plot
ggplot(df, aes(x = condition, y = outcome, color = group, group = group)) +
  geom_point(position = position_dodge(width = 0.2)) +
  geom_line(position = position_dodge(width = 0.2)) +
  labs(x = "Condition", y = "Outcome") +
  scale_color_manual(values = c("blue", "red"),
                     labels = c("AHS", "RT2")) +
  theme_minimal()
```

```{r}
library(emmeans)
emmeans(mod_12cd, list(pairwise ~ group), adjust = "tukey")
```

```{r}
emmeans(Model5, list(pairwise ~ Group), adjust = "bonferroni")
emmeans(mod_11_rtcd, list(pairwise ~), adjust = "bonferroni")
```

# Model estimates for Accuracy

```{r}
model_params_cd <- broom.mixed::tidy(mod_5cd)

model_params_cd |> 
  filter(!str_detect(term, "sd")) |>
  ggplot(aes(x = term, y = estimate)) +
  # geom_hline(yintercept = 1, linetype = "dotted", color = "black") +
  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), width = 0.1, color = "black") +
  geom_point(aes(color = term), size = 3) +
  theme(legend.position = "bottom") +
  labs(y = "Logit", x = "Group") +
  scale_color_viridis_d() +
  coord_flip()

```

# RT (interaction with condition)
tried both log rt and log rt residuals showed mod 6 was a better fit

```{r}

mod_0_rtcd <- lmer(log_rt ~ 1 + (1|participant), 
               data = Gap_final) 

mod_1_rtcd <- lmer(log_rt ~ hs1+hs2+rt1+rt2 + (1|participant), 
               data = Gap_final)

mod_2_rtcd<- lmer(log_rt ~ hs1+hs2+rt1+rt2 + (1|participant) + (1|item), data = Gap_final)

mod_3_rtcd <- lmer(log_rt ~ hs1+hs2+rt1+rt2 * Condition + (1|participant) + (1|item), data = Gap_final)

mod_4_rtcd <- lmer(log_rt ~ length + (hs1 + hs2 + rt1 + rt2) * Condition + (1|participant) + (1|item), data = Gap_final,settingAGQ, control = lmerControl(optimizer= "bobyqa"))

mod_5_rtcd <- lmer(log_rt ~ length + (hs1 + hs2 + rt1 + rt2) * Condition + (1|participant) + (1|item) + (1+condition|participant), data = Gap_final,settingAGQ, control = lmerControl(optimizer= "bobyqa"))

mod_6_rtcd <- lmer(log_rt ~ hs1+hs2+rt1+rt2 + length + Condition + Dom_scale + hs1+hs2+rt1+rt2 * Condition +hs1+hs2+rt1+rt2 * Dom_scale + hs1+hs2+rt1+rt2*Arabic_scale + hs1+hs2+rt1+rt2*English_scale+ (1|participant) + (1|item), data = Gap_final,settingAGQ)

mod_7_rtcd <- lmer(log_rt ~ hs1+hs2+rt1+rt2 + length + hs1+hs2+rt1+rt2 * Condition +hs1+hs2+rt1+rt2 * Dom_scale + hs1+hs2+rt1+rt2*English_scale + hs1+hs2+rt1+rt2*Arabic_scale + (1|participant) + (1|item), data = Gap_final,settingAGQ)



mod_8_rtcd <- lmer(log_rt ~ hs1+hs2+rt1+rt2 + length + Condition + hs1+hs2+rt1+rt2 * Condition +hs1+hs2+rt1+rt2 * Dom_scale + hs1+hs2+rt1+rt2*English_scale + hs1+hs2+rt1+rt2*Arabic_scale + (1|participant) + (1|item), data = Gap_final,settingAGQ, control = lmerControl(optimizer= "bobyqa"))





mod_9_rtcd <- lmer(log_rt ~ length + group + (hs1 + hs2 + rt1 + rt2) * Condition + (hs1 + hs2 + rt1 + rt2) * Arabic_scale + (hs1 + hs2 + rt1 + rt2) * English_scale + (1|participant) + (1|item)  , data = Gap_final,settingAGQ, control = lmerControl(optimizer= "bobyqa"))

mod_10_rtcd <- lmer(log_rt ~ length + group + (hs1 + hs2 + rt1 + rt2) * Condition + (1|participant) + (1|item) + (1+condition|participant), data = Gap_final,settingAGQ, control = lmerControl(optimizer= "bobyqa"))




anova(mod_0_rtcd, mod_1_rtcd, mod_2_rtcd, mod_3_rtcd, mod_4_rtcd, mod_5_rtcd, mod_6_rtcd, mod_7_rtcd, mod_8_rtcd, mod_9_rtcd, mod_10_rtcd)
```

```{r}
Model1 <- glmer(LDTAC_Lists_ABC ~ 1 + Group + Test_Time + Group*Test_Time + (1|Word/Participant_Id) + (Group|Participant_Id), control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=100000)),
```

## I read in Linck and Cunning's paper that, "One rule of thumb is to take t values above 2.0 as being statistically significant". According to the results, it shows that the t values are all below two so there is no main effect of RT (interaction with condition)

Model1 <- glmer(LDTAC_Lists_ABC ~ 1 + Group + Test_Time + Group*Test_Time + (1|Word/Participant_Id) + (Group|Participant_Id), control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=100000)),

```{r}
summary(mod_5_rtcd)
```

```{r}
model_params <- broom.mixed::tidy(mod_5_rtcd)
model_params |>
  filter(!str_detect(term, "sd")) |>
  ggplot(aes(x = term, y = estimate, color = term)) +
  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), width = 0.3, color = "black") +
  geom_point(size = 4) +
  scale_color_manual(values = c("red", "blue", "green", "orange", "purple", "pink", "yellow", "maroon", "black", "brown", "beige", "grey")) +  # Adjust the colors as needed
  theme(legend.position = "bottom") +
  labs(y = "RT", x = "Group")
```

```{r}
anova(mod_6_rtcd)

library(lmerTest)
```

```{r}
plot_model(mod_5_rtcd, type = "diag", show.values = TRUE)
```

```{r}
model_params_rtcd <- broom.mixed::tidy(mod_11_rtcd)

model_params_rtcd |> 
  filter(!str_detect(term, "sd")) |>
  ggplot(aes(x = term, y = estimate)) +
  # geom_hline(yintercept = 1, linetype = "dotted", color = "black") +
  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), width = 0.1, color = "black") +
  geom_point(aes(color = term), size = 3) +
  theme(legend.position = "bottom") +
  scale_color_viridis_d() +
  labs(y = "Estimate", x = "Group") +
  coord_flip()
```
## in this part I tried to follow Flores's Rmd script
# I want to see if all 5 bilingual groups differ about their relative degree of language dominance.
## i started plotting the domincance score, english use and arabic use against group

```{r}
ggplot(data= Gap_final, aes (x=group, y=Dominance_score, fill = group)) +
  geom_bar(stat = "summary")
```

```{r}
ggplot(data= Gap_final, aes (x=group, y=English_use, fill = group)) +
  geom_bar(stat = "summary")
```
```{r}
ggplot(data= Gap_final, aes (x=group, y=Arabic_global, fill = group)) +
  geom_bar(stat = "summary") +
  stat_summary(fun.data = mean_sdl, fun.args = list(mult = 1), 
               geom = "text", aes(label = round(..y.., digits = 2)),
               position = position_stack(vjust = 0.5), size = 4, show.legend = FALSE)
```
```{r}
ggplot(data = Gap_final, aes(x = group, y = English_history, fill = group)) +
  geom_bar(stat = "summary") +
  stat_summary(fun.data = mean_sdl, fun.args = list(mult = 1), 
               geom = "text", aes(label = round(..y.., digits = 2)),
               position = position_stack(vjust = 1.1), size = 4, show.legend = FALSE)
```

```{r}
library(ggplot2)
library(ggplot2)

ggplot(data = Gap_final, aes(x = group, fill = group)) +
  geom_bar(aes(y = Arabic_us), stat = "summary") +
  geom_bar(aes(y = English_use), stat = "summary", position = "dodge") +
  stat_summary(
    fun.data = mean_sdl, fun.args = list(mult = 1),
    geom = "text", aes(label = round(..y.., digits = 2)),
    position = position_stack(vjust = 0.5), size = 4, show.legend = FALSE
  ) +
  facet_grid(. ~ group, scales = "free_x")


```


```{r}
ggplot(data= Gap_final, aes (x=group, y=English_history, fill = group)) +
  geom_bar(stat = "summary")
```

## plot language dominance of 5 bilingual groups as functions of between-language Substractive Differentials, which indicates the direction of dominance.
#the figure shows clearly that AHS, HS2, HS1 are English-dominant, RT1 are Arabic-dominant whereas RT2 are almost balanced bilinguals
```{r}
figure1 <- ggplot (data=Gap_final, aes (x=Dominance_score, y=group, fill =group))+
  geom_boxplot (alpha=0, width=0.3)+
  stat_summary() +
  stat_summary (geom = "errorbar",fun.data = mean_cl_normal, width=0.1)+
  theme_classic() + 
  labs (x= "Between-language Subtractive Differential",
        y= "Bilingual group")
figure1
```
```{r}
figure1 <- ggplot(data = Gap_final, aes(x = Dominance_score, y = group, fill = group)) +
  geom_boxplot(alpha = 0, width = 0.3) +
  stat_summary() +
  stat_summary(geom = "errorbar", fun.data = mean_cl_normal, width = 0.1) +
  theme_classic() +
  labs(x = "Between-language Subtractive Differential",
       y = "Bilingual group") +
  theme(
    axis.title = element_text(size = 16),     # Adjust the font size for axis titles
    axis.text = element_text(size = 14),      # Adjust the font size for axis tick labels
    plot.title = element_text(size = 18)      # Adjust the font size for plot title
  )

figure1
```

```{r}
# Create a custom color palette for your groups
custom_palette <- c("#1f77b4", "#ff7f0e", "#2ca02c", "#d62728", "#9467bd", "#8c564b")

figure1 <- ggplot(data = Gap_final, aes(x = Dominance_score, y = group, fill = group)) +
  geom_boxplot(alpha = 0.7, width = 0.5) +   # Adjust alpha and width as needed
  stat_summary(geom = "point", fun = mean, shape = 18, size = 3) +
  stat_summary(geom = "errorbar", fun.data = mean_cl_normal, width = 0.2) +
  scale_fill_manual(values = custom_palette) +  # Apply the custom color palette
  theme_minimal() +                           # Use a different theme for a cleaner look
  labs(x = "Between-language Subtractive Differential",
       y = "Bilingual group") +
  theme(
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 14),
    plot.title = element_text(size = 18)
  )

figure1
```

```{r}
ggplot(data = Gap_final, aes(x = group, y = English_use, fill = group)) +
  geom_bar(stat = "summary") +
  stat_summary(fun.data = mean_sdl, fun.args = list(mult = 1), 
               geom = "text", aes(label = round(..y.., digits = 2)),
               position = position_stack(vjust = 1.1), size = 4, show.legend = FALSE)
```
```{r}
ggplot(data = Gap_final, aes(x = group, y = Arabic_use, fill = group)) +
  geom_bar(stat = "summary") +
  stat_summary(fun.data = mean_sdl, fun.args = list(mult = 1), 
               geom = "text", aes(label = round(..y.., digits = 2)),
               position = position_stack(vjust = 1.1), size = 4, show.legend = FALSE)
```
```{r}
# Create a single ggplot object
combined_plot <- ggplot(data = Gap_final, aes(x = group)) +
  geom_bar(aes(y = Arabic_use, fill = group), stat = "summary") +
  stat_summary(data = Gap_final, aes(x = group, y = Arabic_use), fun.data = mean_sdl, fun.args = list(mult = 1), 
               geom = "text", aes(label = round(..y.., digits = 2)),
               position = position_stack(vjust = 1.1), size = 4, show.legend = FALSE) +
  geom_bar(aes(y = English_use, fill = group), stat = "summary", position = "dodge") +
  stat_summary(data = Gap_final, aes(x = group, y = English_use), fun.data = mean_sdl, fun.args = list(mult = 1), 
               geom = "text", aes(label = round(..y.., digits = 2)),
               position = position_stack(vjust = 1.1), size = 4, show.legend = FALSE) +
  labs(title = "Arabic and English Use")

# Use facet_wrap to separate by titles
combined_plot <- combined_plot + facet_wrap(~title, scales = "free_x")

# Display the combined plot
print(combined_plot)
```

```{r}
library(ggplot2)
library(gridExtra)

# Your first ggplot code
plot1 <- ggplot(data = Gap_final, aes(x = group, y = English_use, fill = group)) +
  geom_bar(stat = "summary") +
  stat_summary(fun.data = mean_sdl, fun.args = list(mult = 1), 
               geom = "text", aes(label = round(..y.., digits = 2)),
               position = position_stack(vjust = 1.1), size = 4, show.legend = FALSE)

# Your second ggplot code
plot2 <- ggplot(data = Gap_final, aes(x = group, y = Arabic_use, fill = group)) +
  geom_bar(stat = "summary") +
  stat_summary(fun.data = mean_sdl, fun.args = list(mult = 1), 
               geom = "text", aes(label = round(..y.., digits = 2)),
               position = position_stack(vjust = 1.1), size = 4, show.legend = FALSE)

# Combine the two plots side by side
combined_plot <- grid.arrange(plot1, plot2, ncol = 2)

# Print the combined plot
print(combined_plot)
In this code, we first create your two individual ggplot plots (plot1 and plot2) as you provided. Then, we use the grid.arrange function from the gridExtra package to arrange these two plots side by side in a single picture. The ncol argument specifies the number of columns, which in this case is set to 2 to place the plots side by side. Finally, we print the combined_plot to display it. Make sure you have the gridExtra and ggplot2 packages installed in your R environment. You can install them using the install.packages function if you haven't already.





```
```{r}
# Your first ggplot code with Arabic use
plot1 <- ggplot(data = Gap_final, aes(x = group, y = Arabic_use, fill = group)) +
  geom_bar(stat = "summary") +
  stat_summary(fun.data = mean_sdl, fun.args = list(mult = 1), 
               geom = "text", aes(label = round(..y.., digits = 2)),
               position = position_stack(vjust = 1.1), size = 4, show.legend = FALSE) +
  labs(title = "Arabic Use", y = bquote(bold("Arabic Use")))

# Your second ggplot code with English use
plot2 <- ggplot(data = Gap_final, aes(x = group, y = English_use, fill = group)) +
  geom_bar(stat = "summary") +
  stat_summary(fun.data = mean_sdl, fun.args = list(mult = 1), 
               geom = "text", aes(label = round(..y.., digits = 2)),
               position = position_stack(vjust = 1.1), size = 4, show.legend = FALSE) +
  labs(title = "English Use", y = bquote(bold("English Use")))

# Combine the two plots side by side
combined_plot <- grid.arrange(plot1, plot2, ncol = 2)

# Print the combined plot
print(combined_plot)
In this code:

I've added labs(title = "Arabic Use", y = bquote(bold("Arabic Use"))) to the first plot (plot1) to set the title to "Arabic Use" and make the y-axis label bold and black.
Similarly, for the second plot (plot2), I've added labs(title = "English Use", y = bquote(bold("English Use"))) to set the title to "English Use" and make the y-axis label bold and black.
The rest of the code remains the same as in the previous answer to combine the plots side by side and print the combined plot.
Now, your combined plot will have separate titles for Arabic Use and English Use, and the y-axis labels will be in bold and black.





```




```{r}
# Create a single ggplot object
combined_plot <- ggplot(data = Gap_final, aes(x = group)) +
  geom_bar(aes(y = Arabic_use, fill = group), stat = "summary") +
  stat_summary(aes(y = Arabic_use), fun.data = mean_sdl, fun.args = list(mult = 1), 
               geom = "text", aes(label = round(..y.., digits = 2)),
               position = position_stack(vjust = 1.1), size = 4, show.legend = FALSE) +
  geom_bar(aes(y = English_use, fill = group), stat = "summary", position = "dodge") +
  stat_summary(aes(y = English_use), fun.data = mean_sdl, fun.args = list(mult = 1), 
               geom = "text", aes(label = round(..y.., digits = 2)),
               position = position_stack(vjust = 1.1), size = 4, show.legend = FALSE) +
  labs(title = "Arabic and English Use")

# Use facet_wrap to separate by titles
combined_plot <- combined_plot + facet_wrap(~title, scales = "free_x")

# Display the combined plot
print(combined_plot)
```


```{r}
# Calculate the required summary statistics
summary_stats <- Gap_final %>%
  group_by(group) %>%
  summarise(
    mean_diff = mean(Dominance_score),
    min_diff = min(Dominance_score),
    max_diff = max(Dominance_score),
    sd_diff = sd(Dominance_score)
  )
print(summary_stats)
```
```{r}
install.packages("gamlss")

library(gamlss)

ModelQ1 <- gamlss(Dominance_score ~ group, family = BE, data = na.omit(Gap_final))
summary(ModelQ1)
confint(ModelQ1)
```



#Run a regression model to confirm the hypothesis that whether RT2 is more balanced than the other groups.
#This time, the outcome variable is the Dominance Index (ratio) proposed by Birdsong (2016). This measure was chosen because it describes the language balance as a continuous variable. An one-inflated beta-regression was run because the outcome varies between (0, 1]. 

```{r}
Gap_final$Dominance_score <- scale(Gap_final$Dominance_score)
colnames(Gap_final)[35] <- "scale_dominance"
```

```{r}
hist(Gap_final$scale_dominance)
log_dominance <- log(Gap_final$scale_dominance)
hist(log_dominance)
```
## didn't converge when added item showed error

```{r}
lmer(Dominance_score ~ group + (1|participant), data= Gap_final)
```
x

```{r}
summary (Model_dom)
```

```{r}
# Conduct a one-way ANOVA
myanova <- aov(English_use ~ group, data = Gap_final)

# Check the summary of the ANOVA results
summary(myanova)

# Conduct a post-hoc Tukey test
mytukey <- TukeyHSD(myanova)

# Check the results of the Tukey test
mytukey
```

```{r}
hist(Gap_final$English_use, main = "Histogram of Data")
```
```{r}
spearman_result <- cor(Gap_final$gap_acc_cong, Gap_final$English_use, method = "spearman")
```

```{r}
subset_dataAHS <- Gap_final[Gap_final$AHS == "AHS", ]
subset_dataRT1 <- Gap_final[Gap_final$RT1 == "RT1", ]
subset_dataRT2 <- Gap_final[Gap_final$RT2 == "RT2", ]

# Calculate the correlation within the subgroup
correlation_result <- cor(subset_data$lexical_decision_performance, subset_data$english_use, method = "pearson")

# Print the correlation result
print(correlation_result)
```


